\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{enumitem} % Required for custom enumerate labels

\title{Questions}
% \subtitle{UENG201-Introduction to Materials Science}
% The article class does not support the subtitle command. You can include this information in the title or as part of the document content.
\author{Animesh Pradhan}

\begin{document}
\maketitle
\section{Interstellar Radio}
An astronomer looking for faint stars finds a strong and consistent radio signal from a very distant object on the ecliptic and tracks the source for four years. By analysing the ecliptic longitude data, he finds that its time dependence can be represented by a linear term and a periodic term, with a one year period. The total linear drift in the first year is $52.3''$. The same value in the fourth year(total drift for that year) is $52.62''$. The periodic term is a sinusoidal function of the ecliptic longitude of the sun with respect to the object, $\lambda_{rel}$. When expressed as a sum of sine and cosine terms with no phase shift, the sin term has an amplitude of $0.500''$ in the first year and $0.580''$ in the fourth year. Whereas the cosine term has an amplitude of $20.49''$ in both years. Take the general rate of precession to be a constant at $50.30''/yr$. State and explain all the effects which lead to these variations and find the ratio of the distance of the object from the solar system in Year 4 compared to Year 1. Also, find the ratio of apparent angular velocities in Year 4 compared to Year 1. Finally, use these ratios to conclude the nature of the object's motion. Assume radial motion of the object, if any, remains constant throughout the observation period.

\subsection{Brief Solution}
Total linear drift is the sum of that due to precession and proper motion. Since the precession rate is constant, the apparent angular velocity changes with time. The sin term corresponds to the parallax which increases, implying decreasing distance, while the cosine term corresponding to stellar aberration remains constant, confirming earth's motion to not be the cause of irregularities.  The ratio of the distance is simply inverse of the ratio of parallax, $0.500/0.580 = 0.862$ .
The apparent angular velocities are $52.30''/yr - 50.30''/yr = 2.00''/yr$ and $52.62''/yr - 50.30''/yr = 2.32''/yr$. The ratio of these velocities is the same, at $0.862$. This indicates that the transverse velocities are the same for both year. Constant radial velocities imply that the object is unbounded to the sun, however, is moving towards the solar system at a constant rate.

\section{Black Holes giveth and Black Holes taketh away}
The growth of a supermassive black hole (SMBH) in a galaxy with a finite gas reservoir of total mass $M_{res}$ can be modeled as a self-limiting process. While mass continually accretes into the SMBH, powerful jets are ejected perpendicular to the accretion disk. The effective mass accretion rate, $\dot{M}_{acc}$, is described by the logistic equation:
\[
\dot{M}_{acc}(t) = \lambda M_{BH}(t) \left(1 - \frac{M_{BH}(t)}{M_{res}}\right)
\]
where $M_{BH}(t)$ is the black hole's current mass and $\lambda$ is a constant representing the initial growth rate. The fraction of the accreted mass-energy converted into both jets combined have a kinetic power given by $P_{jet} = \epsilon \dot{M}_{acc} c^2$, where $\epsilon$ is the jet efficiency. Assume mass of SMBH at time $t=0$ is $M_0$.

\begin{enumerate}
    \item Find the mass of the black hole at a given time $t$
    \item Derive an expression for the total kinetic energy, $E_{total}$, that a single SMBH will ever inject into its host galaxy, in the form of jets, over its entire lifetime (from $t=0$ to $t=\infty$). Express your answer in terms of the energy efficiency $\epsilon$, the speed of light $c$, and the masses $M_0$ and $M_{res}$.

    \vspace{1em}

    \item Now, consider a large cluster containing $N$ such galaxies. Black hole formation in the cluster was a continuous process, resulting in a population of SMBHs with ages uniformly distributed between $t=0$ and a maximum age $T$. Derive an expression for the total, instantaneous kinetic power, $P_{total}$, injected by the combined jets of this entire population. Express your answer in terms of $N, \epsilon, c, T$, and the masses $M_{BH}(T)$ (the mass of a black hole at age T) and $M_0$. Also, in the form of an implicit equation, derive the maximum age $T_{max}$ for which the total power is maximized, $P_{total}(T_{max})$.
\end{enumerate}

\subsection{Brief Solution}
\begin{enumerate}
    \item Integrating the logistic equation we get $M_{BH}(t) = \dfrac{M_{res}}{1+A e^{-\lambda t}}$, where $A = \dfrac{M_{res}}{M_0} - 1$
    \item $ E_{total} = \int_0^\infty P_{jet}(t) \,dt = \int_0^\infty \epsilon \dot{M}_{acc}c^2 \,dt = \epsilon c^2 \int_0^\infty \frac{dM_{BH}}{dt} dt = \epsilon c^2 \int_{M_{BH}(0)}^{M_{BH}(\infty)} dM_{BH}$
        \\ $ = \epsilon c^2 (M_{BH}(\infty) - M_{BH}(0)) = \epsilon c^2 (M_{res} - M_0)$
    \item $ P_{total}(t) = \int_{0}^{T} \dfrac{N \epsilon c^2}{T} \dfrac{dM_{BH}}{dt} dt = \dfrac{N \epsilon c^2}{T} \int_{0}^{T} \dfrac{dM_{BH}}{dt} dt = \dfrac{N \epsilon c^2}{T} (M_{BH}(T) - M_{BH}(0))$
          \\ $\dfrac{dP_{total}}{dT} = 0 \Rightarrow T_{max} \dot{M_{acc}}(T_{max}) = M_{BH}(T_{max}) - M_{BH}(0)$. This states that the total power is maximized when the instantaneous accretion rate of the oldest black holes is exactly equal to their average rate over their entire lifetimes.
\end{enumerate}
\end{document}